---
title: "ch_1_introduction_notebook"
author: "Jason Grahn"
date: "April 8, 2018"
output: html_notebook
---

#Script 1.1 R-as-a-calculator

For the simple commands for math in R, I'll use the same code, but commenting it out for practice. 
```{r}
#simple addition
1+1

#multiplication and exponents including order of operations
5*(4-1)^2

#square roots
sqrt(10)

#logs
log(10)

#square roots AND logs
sqrt(log(10))
```

#Script 1.2 installs a bunch of packages

This R script downloads and installs all packages used at some point. It needs to be run once for each computer/user only. The list in the book has 28 packages involved and I'm going to add the tidyverse (and maybe others to it as well.)

We don't require the output, so I'm adding `echo = FALSE` to the code block header.

```{r install packages}
#make a vector of all the known-needed packages
package_array <- (c("AER", "car", "censReg", "dplyr", "dummies", "dynlm", "effects", "ggplot2",
                    "lmtest", "maps", "mfx", "orcutt", "plm", "quantmod", "sandwich", "quantreg", 
                    "rio", "rmarkdown", "sampleSelection", "stargazer", "survival", "systemfit",
                    "truncreg", "tseries", "urca", "xtable", "vars", "WDI", "xts", "zoo", 
                    "wooldridge", "tidyverse", "tibble"))

#install all the packages in that vector
#I'm commenting out the install packages code because I dont want to install these packages everytime I knit this file!
#install.packages(package_array, repos = "http://cran.us.r-project.org")

#load all those packages at once
lapply(package_array, require, character.only = TRUE)
```


#Script 1.3 is Making objects

This is other basic R work. Objects are assigned by using `<-`. I dont know that tidy improves on this at all, so going to just run with what Heiss input with my own code comments

```{r}
#make something the object *x*, this won't have any output.
x <- 5

#display x, and x^2
x
x^2

#make objects y & z with immediate display
#to make something immediate display put the command in *()*
(y <- 3)

#z is y to the x power
(z <- y^x)

```

skipping ahead to..

#Script 1.8 about Naming and Index Vectors

```{r Script 1.8.a Naming and Index Vectors}
#create a vector of averages #this is directly from the book
avgs <- c(.336, .358, .356, .349, .346)

#create a vector of player names
players <- c("Cobb","Hornsby", "Jackson", "o'Doul", "Delahanty")

#assign names to vectors and display vectors
names(avgs) <- players
avgs

#extracting just one indices by number
avgs[2]

#or first through fouth indices
avgs[1:4]

#or look up an indices by name
#avgs["jackson"] This doesn't work because 'Jackson' is upper case
avgs["Jackson"]

#and we can do logical indices
avgs[avgs >= 0.35]
```


Let's remake that vector in the tidy way. Each player name and average will be an observation. We'll need tibble for this.

```{r Script 1.8.b Naming and Index Vectors}
library(tibble)
player_array <- tibble(x = players, y = avgs)
player_array

#that's good, but those need headers. 
player_array_with_headers <- tibble(player_name = players, 
                                    player_avgs = avgs)

player_array_with_headers

#ok, so how do I call an individual observation row?  
#because `player_array_with_headers[2]` calls the entire second column...
#ah! call the second observation (row) as the first argument inside [], with nothing as the second argument. 
player_array_with_headers[2,]

#so if we want the first through fourth observation
player_array_with_headers[1:4,]

#and now how about viewing just one observation? I'd do that using the pipe and a dplyr filter
#need to load dplyr that was installed with the tidyverse
library(dplyr)
player_array_with_headers %>% 
  filter(player_name == "Jackson")
```

skipping ahead...

#Script 1.12 about data frames part 1, making a few frames to be used in 1.13


```{r Script 1.12 Data Frames}
#define a matrix of x values:
year <- c(2008, 2009, 2010, 2011, 2012, 2013)

#define a matrix of y values:
#the ";" has the same effect as making a new line. this is just compacting the code.
product1 <- c(0,3,6,9,7,8); product2 <- c(1,2,3,5,9,6); product3 <- c(2,4,4,2,3,2)

#bind those columns to a matrix
sales_mat <- cbind(product1, product2, product3)

#name the rows using the year matrix
#this isn't a best practice according to the tibble information.
rownames(sales_mat) <- year

#gives us the sales_mat matrix with named rows and columns
sales_mat

#now we're making it a data.frame
sales <- as.data.frame(sales_mat)

sales
```

# Now onto 1.13a, working with data frames.

```{r Script 1.13.a Data Frames}
#accessing a single variable
sales$product2

#generating a new variable in the dataframe
sales$totalv1 <- sales$product1 + sales$product2 + sales$product3

#same using "with"
sales$totalv2 <- with(sales, product1 + product2 + product3)

#or using "attach"
#which seems weird, because "attach" is for databases?
attach(sales)
sales$totalv3 <- product1 + product2 + product3 
detach(sales)
sales
#that's too many steps imo
```



#1.13b is me trying the same with tidy commands\

```{r Script 1.13.b Data Frames}
#first lets remake that as a tibble 
sales_1_13b <- tibble(year = year,
                      product1 = product1,
                      product2 = product2,
                      product3 = product3)
sales_1_13b

#now let's make that extra column
sales_1_13b <- sales_1_13b %>% 
  mutate(totalv1 = product1 + product2 + product3)
sales_1_13b

#comparing mutate to the options in the book, the only one that makes sense for small datasets (to me) is 
#sales$totalv1 <- sales$product1 + sales$product2 + sales$product3
#but that seems to have a bunch of redundant coding with the "sales$" being reused so often
```

#Subsets of data

```{r 1.14a and b since a small section}
sales
#subset the sales table to show years where sales of product3 were >= 3
subset(sales, product3>=3)

#wonder if i can use %>% with subset?
sales %>% 
  subset(product3 >= 3)
#yep.

#Alternatively, we can filter.
sales %>% 
  filter(product3 >= 3)
#but doing this loess the rowname. interesting.
#ah, this is/was a known issue https://github.com/tidyverse/dplyr/issues/366
#Hadley "Yes, all dplyr methods ignore rownames.... On purpose. They're hard to get right in general 
#(esp because of uniqueness criterion) and don't allow you to do anything that you can't otherwise do."
#so i guess don't use rownames.

#if the data is tidy then... 
sales_1_13b %>% 
  filter(product3 >= 3)
#has the same result as...
sales_1_13b %>% 
  subset(product3 >= 3)

```

#giving a pause here for now. page 22, 1.3.3 "R data files"
